Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method

### 2017-04-15 doc/M&C_2017/Presentation/MC2017_Presentation.tex
Presentation

\setbeamertemplate{footline}{
\leavevmode%
%\hbox{\hspace*{-0.06cm}
\begin{beamercolorbox}[wd=.5\paperwidth,ht=3.25ex,dp=1ex,center]{author in head/foot}%
	\usebeamerfont{author in head/foot}\insertshortauthor%~~(\insertshortinstitute)
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.25\paperwidth,ht=3.25ex,dp=1ex,center]{section in head/foot}%
	\usebeamerfont{section in head/foot} Final Exam March 2017 % \insertshorttitle
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.25\paperwidth,ht=3.25ex,dp=1ex,left]{section in head/foot}%
	\usebeamerfont{section in head/foot}\insertshortdate{}\hspace*{2em}
	\insertframenumber{} / \inserttotalframenumber %\hspace*{2ex}
\end{beamercolorbox}}%
%\vskip0pt%
%}

@@

\title{Improved Quasi-Static Methods for Time-Dependent Neutron Diffusion and Implementation in Rattlesnake}

\author[Zachary M. Prince]{\Large Zachary M. Prince \\ \vspace{5mm} \normalsize \underline{Committee} \\ Dr. Jean Ragusa (Chair) \\ Dr. Jim Morel \\ Dr. Bojan Popov}
\institute{Final Examination for Partial Fulfillment of a Masters of Science \\ Department of Nuclear Engineering, Texas A\&M University, College Station, TX}

\logo{\includegraphics[height=0.8cm]{figures/TAM-logo.png}}

@@

%-------------------------------------------------------------------
\begin{frame}{Objective}

\begin{block}{Purpose Statement}
Investigate and develop the improved quasi-static method (IQS) for minimizing computation expense in transient reactor simulations, while maintaining accuracy.
\end{block}

\begin{block}{What is IQS?}
\begin{itemize}
\item Transient neutronics method meant to minimize expensive diffusion evaluations
\item Non-approximately separates spatial variance and temporal power
\item Impetus is spatial variance doesn't change much in time, compared to powerb
\end{itemize}
\end{block}

\begin{block}{Objectives}
\begin{itemize}
\item Establish IQS performance for various nonlinear iteration techniques applied to transient neutronics
\item Validate time step convergence in transient neutronics problems for IQS with high order discretization schemes
\item Apply IQS to multiphysics simulations
\end{itemize}
\end{block}

%\begin{block}{Conclusions}
%\begin{itemize}
%\item IQS is an effective method for reducing computational expense of transient reactor simulations
%\item The IQS uniqueness criteria is important to keep consistent through nonlinear iteration and accurate precursor evaluation
%\item IQS demonstrates proper error convergence up through fourth order methods and shows impressive performance with time adaptation
%\item IQS implementation in Rattlesnake was effective and greatly reduced computational expense in TREAT simulations
%\end{itemize}
%\end{block}

@@

%-------------------------------------------------------------------
\begin{frame}{Transient Testing}

\begin{block}{}
\begin{itemize}
\item Fukushima sparked a systematic demand for accident tolerant reactor system designs
\item New designs are significantly different and require experimental testing 
\item Tesing requires specialized facilities like ACRR at SNL and TREAT at INL
\end{itemize}
\end{block}

\begin{figure}
\includegraphics[height=1.5in]{figures/fukushima_melt.jpg}
\caption{Hole believed to be caused by fuel leaking out of the pressure vessel at Fukushima Daiichi nuclear power plant}
\end{figure}

@@

%------------------------------------------------------------------%
\subsection{Transient Reactor Simulation}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{Modeling and experimentation are mutualistic}

\begin{figure}
\includegraphics[width=\linewidth]{figures/relationship_chart.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{Transient reactor simulation is difficult}

\begin{block}{Neutronics}
\begin{itemize}
\item Transient behavior of neutrons is typically evaluated using deterministic neutron transport to determine flux
\item Neutron transport contains 7 independent variables: time ($t$), position ($\vr$), energy ($E$), direction ($\vo$)
\item Neutron diffusion carefully eliminates direction to reduce to 5 variables
\item These equations result in a stiff system, so expensive implicit solvers are required for stability
\item Each variable needs to be discretized: FEM for space, multigroup for energy, and time will be discussed later
\end{itemize}
\end{block}

\begin{block}{Multiphysics}
\begin{itemize}
\item Reactor physics involve more than the behavior of neutrons, including: fuel temperature, thermal hydraulics, fuel structure, depletion etc.
\item All of these variables are dependent on one another, or coupled
\item Each variable have very different solution methods
\item The exponential increase in computational expense for each variable and communication between different solvers make multiphysics a daunting task
\end{itemize}
\end{block}

@@

%------------------------------------------------------------------%
\subsection{Point Reactor Kinetics}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{Point Reactor Kinetics Equation (PRKE)}

\begin{block}{Factorization}
Point reactor kinetics involve factorizing flux into space and time dependent components:
\be
\phi^g(\vr,t) = \Phi^g(\vr)\times p(t)
\ee
\end{block}

\begin{block}{PRKE}
\[
\frac{dp}{dt}=\left[\frac{\rho-\bar{\beta}}{\Lambda}\right]p+\sum_{i=1}^I\bar{\lambda}_i\xi_i
\]
\[
\frac{d\xi_i}{dt}=\frac{\bar{\beta}_i}{\Lambda}p - \bar{\lambda}_i\xi_i \quad 1 \le i \le I 
\]
\end{block}

\begin{block}{}
\begin{itemize}
\item Large assumption of time-independent spatial variance
\item Very inexpensive to evaluate after initial flux calculation
\item "Back-of-the-envelope" calculation and postprocessing
\end{itemize}
\end{block}

@@

%\begin{block}{Variable time discretization}
%\begin{itemize}
%\item To ensure stability, only implicit discretization schemes
%\item PRKE error kept constant, so either very small time steps or adaptive methods
%\item Schemes for shape include:
%\begin{itemize}
%\item Implicit Euler
%\item Crank-Nicolson
%\item Backward Difference Formulae (DDF)
%\item Singly-Diagonally-Implicit Runge-Kutta (SDIRK)
%\end{itemize}
%\end{itemize}
%\end{block}

%-------------------------------------------------------------------
\begin{frame}{Interpolation of PRKE Parameters}

\begin{block}{Lagrange Interpolation}
\begin{itemize}
\item Parameters are evaluated at each macro step using the relevant \tcb{shape}
\item Parameters are interpolated between macro steps for PRKE
\item Most applications use linear interpolation
\item Higher order Lagrange is possible:
\be
\rho(t) = \sum_{j=0}^k \rho_{n-j+1} \prod_{m\neq j}^k \frac{t-t_{n-m+1}}{t_{n-j+1}-t_{n-m+1}}
\ee
\end{itemize}
\end{block}

\begin{figure}
\includegraphics[height=1.5in]{figures/param_interp.png}
\end{figure}

@@

%-------------------------------------------------------------------
\begin{frame}{Relevant Picard convergence criteria}

\begin{block}{IQS convergence criteria}
\begin{itemize}
\item \tcb{Shape} based convergence:
\be
\frac{\text{max}\left|\tcb{\varphi{n}^{(k+1)}} - \tcb{\varphi_{n}^{(k)}}\right|}{\text{max}\left|\tcb{\varphi{n}^{(k+1)}}\right|} < \epsilon_{\varphi}
\ee 
\be
\frac{\norm{\tcb{\varphi{n}^{(k+1)}} - \tcb{\varphi_{n}^{(k)}}}}{\norm{\tcb{\varphi_{n}^{(k+1)}}}} < \epsilon_{\varphi}
\ee

\item Reactivity based:
\be
\left(\frac{\rho}{\Lambda}\right)^{(k+1)} - \left(\frac{\rho}{\Lambda}\right)^{(k)} < \epsilon_{\rho}
\ee

\item \tcr{Amplitude} based:
\be 
\tcr{p_n^{k+1}} - \tcr{p_n^{k}} < \epsilon_p
\ee

\item Uniqueness consistency:
\be 
\frac{K_n^{(k+1)} - K_0}{K_0} < \epsilon_K
\ee
\end{itemize}
\end{block}

@@

%------------------------------------------------------------------%
\subsection{Time Discretization}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{Time Discretization}
\vspace{-3mm}

\begin{block}{Time Discretizations}
\begin{itemize}
\item Implicit Euler - 1st Order
\item Crank-Nicolson - 2nd Order
\item Backward Difference Formulae (BDF) - 1st, 2nd, 3rd, 4th Order
\item Singly-Diagonally-Implicit Runge-Kutta (SDIRK) - 3rd Order (SDIRK33)
\end{itemize}
\end{block}

\begin{block}{Time Step Adaptation}
\begin{itemize}
\item Determining appropriate time step is difficult without prior knowledge of the solution
\item Some areas of the transient may require smaller time steps for constant accuracy
\item Time adaptation predicts the local truncation error and computes a time step appropriate for a given tolerance
\item Examples:
\begin{itemize}
\item Step doubling
\item Embedded Runge-Kutta
\end{itemize}
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Step doubling time adaption was chosen for testing}

\begin{block}{Step Doubling Solution Process}

\begin{figure}[htpb!]
\centering
\begin{tikzpicture}[scale=1.5]
\draw[] (0,2) -- (6,2) ;
\foreach \x in  {0,3,6}
\draw[shift={(\x,2)},color=black] (0pt,6pt) -- (0pt,-6pt);
\draw[shift={(0,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_n$};
\draw[shift={(3,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1/2}$};
\draw[shift={(6,2)},color=black] (0pt,0pt) -- (0pt,-6pt) node[below] {$t_{n+1}$};
\draw (0,2) edge[out=45,in=135,->,thick,blue] node[above,sloped] {\tcb{Step 1}} (6,2);
\draw (0,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 2}} (3,2);
\draw (3,2) edge[out=25,in=155,->,thick,red] node[above,sloped] {\tcr{Step 3}} (6,2);
\node[anchor=west](shape) at (6.1,2.5) {$\tcb{\phi_{\Delta t}^g}$};
\node[anchor=west](shape) at (6.1,2) {$\tcr{\phi_{\Delta t/2}^g}$};

\end{tikzpicture}
\end{figure}
\be
e_n =\frac{\norm{\sum_{g=1}^G \left(\tcr{\phi^g_{\Delta t/2}} - \tcb{\phi^g_{\Delta t}} \right)}}{\max\left(\norm{\sum_{g=1}^G\tcr{\phi^g_{\Delta t/2}}},\norm{\sum_{g=1}^G\tcb{\phi^g_{\Delta t}}}\right)} 
\qq  \qq 
\Delta t_{new} = S_f \Delta t \left[\frac{e_{tol}}{e_n}\right]^{1/(p+1)}   
\ee

%-------------------------------------------------------------------
\begin{frame}{Step Doubling Solution Process}

\begin{block}{Programming Visualization}

\begin{figure}[!htpb]
\centering
\begin{tikzpicture}[every node/.style = {font=\small},scale=0.5]

\node[blueblock](p1) at (-6,3) {Step 1} ;
\node[redblock](p2) at (-8,0) {Step 2} ;
\node[redblock](p3) at (-4,0) {Step 3} ;
\node[purpleblock](p4) at (0,0) {Compute \\ $e_n$};
\node[purpleblock](p5) at (4,0) {Compute  \\ $\Delta t_{new}$};
\node[greendiamond](p6) at (8,0) {$e_n < e_{tol}$};
\node[orangeblock](p7) at (0,-3) {Restore \\ Solutions};

\draw[->,thick](p1.east) -| (p4.north);
\draw[->,thick](p2.east) -- (p3.west);
\draw[->,thick](p3.east) -- (p4.west);
\draw[->,thick](p4.east) -- (p5.west);
\draw[->,thick](p5.east) -- (p6.west);
\draw[->,thick](p6.south) |- node[right] {no} (p7.east);
\draw[->,thick](p6.east) -- node[above] {yes} (11,0);
\draw[->,thick](p7.west) -| (-10,0) -- (p2.west);
\draw[->,thick](p7.west) -| (-10,0) |- (p1.west);

\end{tikzpicture}
\end{figure}
Each Step undergoes:
\bi
\item Shape evaluation
\item PRKE evaluations
\item Multiphysics evaluations
\item Iterations for convergence of amplitude, shape, and multiphysics
\ei

%------------------------------------------------------------------%
\subsection{Delayed Neutron Precursors}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{Treatment of delayed neutron precursors}
\vspace{-3mm}
\begin{block}{Delayed neutron precursor equation is an ODE}
\begin{equation*}
\frac{dC_i}{dt} = \tcr{p}\sum_{g=1}^G \beta_i\nu \Sigma_f^g \tcb{\varphi^{g}} - \lambda_iC_i , \quad 1 \le i \le I
\end{equation*}
\end{block}

\begin{block}{Time Integration of Precursors}
\begin{itemize}
\item Trapezoid Rule:
\be
C^{n+1}_i = \frac{1-0.5\Delta t\lambda_i}{1+0.5\Delta t\lambda_i}C^n_i + \frac{0.5\Delta t \beta_i}{1+0.5\Delta t\lambda_i}\sum_{g=1}^G\left[(\nu\Sigma_f^g\tcb{\varphi^g})^n \tcr{p^n} +  (\nu\Sigma_f^g\tcb{\varphi^g})^{n+1} \tcr{p^{n+1}}\right]
\ee
\item Analytical Integration:
\be 
C^{n+1}_i = C^n_i e^{-\lambda_i \Delta t} + \sum_{g=1}^G\left(\hat{a}_{2,i} (\nu\Sigma_f^g\tcb{\varphi^g})^{n+1}+\hat{a}_{1,i} (\nu\Sigma_f^g\tcb{\varphi^g})^n\right)\beta_i
\ee
\vspace{-4mm}
\begin{align*}
&\hat{a}_{1,i}= \int_{t_n}^{t_{n+1}}\frac{t_{n+1}-t'}{\Delta t}\tcr{p(t')}e^{-\lambda_i(t_{n+1}-t')}dt' \\
&\hat{a}_{2,i} = \int_{t_n}^{t_{n+1}}\frac{t'-t_n}{\Delta t}\tcr{p(t')}e^{-\lambda_i(t_{n+1}-t')}dt'
\end{align*}
\end{itemize}
\end{block}

@@

%------------------------------------------------------------------%
\subsection{MATLAB Prototype}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{MATLAB Kinetics Prototype Code}

\begin{block}{Overview}
\begin{itemize}
\item One-dimensional, single-group, time-dependent neutron diffusion with precursors
\item Heterogenous medium using block-like regions
\item Step and ramp perturbations in parameters
\item Manufactured solutions using symbolic capability
\item FEM meshing with any order shape functions
\end{itemize}
\end{block}

\begin{block}{Solution Methods}
\begin{multicols}{2}
\begin{enumerate}
\item Implicit Discretization
\begin{itemize}
\item Precursor variable coupling
\item Theta precursor elimination
\item Analytical precursor elimination
\end{itemize}
\item IQS
\begin{itemize}
\item Precursor variable coupling
\item Theta precursor elimination
\item Analytical precursor elimination
\end{itemize}
\item IQS P-C
\begin{itemize}
\item Theta precursor elimination
\item Analytical precursor elimination
\end{itemize}
\end{enumerate}
\end{multicols}
\end{block}

@@

%------------------------------------------------------------------%
\subsection{One-Dimensional Slab}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{One Dimensional Example}

\begin{block}{Geometry}
\begin{figure}
\begin{center}
\resizebox{\linewidth}{!}{
\begin{tabular}{| l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l | l |}
\hline \hline \hline
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
1 & 1 & 1 & 1 & 2 & 3 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 1 & 4 & 4 & 1 & 1 & 1 & 1 \\
  &   &   &   &   &   &    &    &   &   &   &   &   &   &   &   &   &   &   &   \\
\hline \hline \hline
\end{tabular}}
\end{center}
\end{figure}
\begin{itemize}
\item Region 2: Ramp $\Sigma_a$ decrease (rod removal)
\item Region 3: Ramp $\Sigma_a$ decrease then increase (rod removal then insertion)
\item Region 4: Ramp $\Sigma_a$ increase (rod insertion)
\end{itemize}
\end{block}

\begin{block}{Region Perturbations}
\centering
\begin{tabular}{llllllll}
\hline
Region & Material Property & 0.0 s & 0.1 s & 0.6 s & 1.0 s & 1.7 s \\
\hline
2 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.095 & 1.095 & 1.095 \\
3 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.09 & 1.09 & 1.1 \\
4 & $\Sigma_{a} (cm^{-1})$ & 1.1 & 1.1 & 1.105 & 1.105 & 1.105 \\
\hline
\end{tabular}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Baseline Results}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.45\linewidth]{figures/1D_flux.png}
\includegraphics[width=0.55\linewidth]{figures/1D_power_base.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{IQS Results}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.45\linewidth]{figures/1D_shape2.png}
\includegraphics[width=0.55\linewidth]{figures/1D_iqs_power.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{Convergence criteria convergence behavior}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_renorm.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{IQS converged error with theta integration of precursors}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_error.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{IQS converged error with analytical integration of precursors}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/iter_error_an.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{First and second order time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_IE.png}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF2.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{Third and fourth order time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF3.png}
\includegraphics[width=0.5\linewidth]{figures/1D_conv_BDF4.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{SDIRK33 time step error convergence}

\begin{figure}
\includegraphics[width=\linewidth,height=3in,keepaspectratio]{figures/1D_conv_SDIRK33.png}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{Rattlesnake time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\paperwidth]{figures/1D_conv_Rat.png}
\includegraphics[width=0.5\paperwidth]{figures/1D_conv_lin.png}
\end{figure}
\end{columns}

%------------------------------------------------------------------%
\subsection{TWIGL Benchmark}
%------------------------------------------------------------------%

%-------------------------------------------------------------------
\begin{frame}{TWIGL Benchmark}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/twigl_geom.png}
\includegraphics[width=0.5\linewidth]{figures/ndiff_ramp_flux.jpg}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{TWIGL Power Profile}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot.png}
\caption{TWIGL Power Profile}
\end{figure}

\column{.55\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot2.png}
\caption{TWIGL Peak Power Profile}
\end{figure}

%-------------------------------------------------------------------
\begin{frame}{TWIGL time step error convergence}

\begin{columns}
\column{\dimexpr\paperwidth-10pt}
\begin{figure}
\includegraphics[width=0.5\linewidth]{figures/TWIGL_convergence.jpg}
\includegraphics[width=0.5\linewidth]{figures/TWIGL_ramp_lin.png}
\end{figure}
\end{columns}

%-------------------------------------------------------------------
\begin{frame}{TWIGL with Time Adaptation}

\column{.7\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/TWIGL_power_plot_dt2.png}
\caption{TWIGL Power Profile}
\end{figure}

\column{.4\textwidth}
\begin{table}
\begin{center}
\colorbox{white}{\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|}
\hline
\multicolumn{5}{|c|}{Implicit Discretization}\\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.00012677 &	9    &	29 \\
2 &	0.01    &	3.5555e-05 &	11   &	35 \\
3 &	0.005   &	4.0364e-05 &	11   &	31  \\
4 &	0.001   &	0.00294822 &	33   &	122  \\
5 &	0.0005  &	0.00099778 &	39   &	131  \\
\hline
\rowcolor{yellow} 6 &	0.0001  &	0.00019510 &	78   &	236  \\
\hline
7 &	5.0e-05 &	0.00018372 &	112  &	342  \\
8 &	1.0e-05 &	8.0564e-05 &	263  &	794 \\
\hline

\multicolumn{5}{|c|}{IQS} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05 	&	0.03380433 &	4   &	20   \\
2 &	0.01 	&	0.00166991 &	5   &	40  \\
3 &	0.005  	&	0.00886584 &	5   &	40  \\
4 &	0.001 	&	0.02976305 &	5   &	36  \\
5 &	0.0005 	&	0.00143781 &	6   &	55  \\
\hline
\rowcolor{yellow} 6 &	0.0001 	&	0.00016175 &	8   &	65  \\
\hline
7 &	5.0e-05 &	6.0328e-05 &	12  &	163  \\
8 &	1.0e-05 &	7.7103e-05 &	379 &	5729 \\
\hline

\multicolumn{5}{|c|}{IQS P-C} \\
\hline
Test & $e_{tol}$ & Error & Steps & Solves \\
\hline
1 &	0.05    &	0.03380433 &	4  &	9   \\
2 &	0.01    &	0.00263068 &	5  &	12  \\
3 &	0.005   &	0.00160486 &	6  &	21  \\
\hline
\rowcolor{yellow} 4 &	0.001   &	1.7527e-05 &	10 &	35  \\
\hline
5 &	0.0005  &	1.4185e-05 &	16 &	74  \\
6 &	0.0001  &	6.2903e-06 &	19 &	78  \\
7 &	5.0e-05 &	1.5247e-06 &	24 &	92  \\
8 &	1.0e-05  &	9.8321e-07 &	48 &	210 \\
\hline
\end{tabular}}}
\end{center}
\end{table}

@@

%-------------------------------------------------------------------
\begin{frame}{LRA with Time Adaptation}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/LRA_DT2.png}
\caption{LRA Power Profile}
\end{figure}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/LRA_DT2_peak.png}
\caption{LRA Peak Power Profile}
\end{figure}

\begin{table}
\begin{center}
\resizebox{\textwidth}{!}{
\begin{tabular}{|l|l|l|l|l|l|l|}
\hline
 & \multicolumn{3}{|c|}{Brute Force} & \multicolumn{3}{|c|}{IQS P-C} \\
\hline
Event & Power (W/cm$^3$) & Error & Steps & Power (W/cm$^3$) & Error & Steps \\
\hline
Max Power & 5567.3 & 0.019454 & 423 & 5568.3 & 0.019274 & 47 \\
End (3 s) & 109.66 & 2.3650e-4 & 603 & 109.65 & 3.0622e-4 & 97 \\
\hline
\end{tabular}}
\end{center}
\end{table}

@@

%-------------------------------------------------------------------
\begin{frame}{Transient-15 with Time Adaptation}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_DT2.png}
\caption{TREAT Power Profile}
\end{figure}

\column{.6\textwidth}
\begin{figure}
\includegraphics[width=\linewidth]{figures/Tran15_DT2_peak.png}
\caption{TREAT Peak Power Profile}
\end{figure}

@@

%-------------------------------------------------------------------
\begin{frame}{Conclusions}

\begin{block}{IQS Iteration}
IQS was testing with fixed point itertion used the following criteria:
\begin{enumerate}
\item $L^\infty$ norm of the change in shape between iterations
\item $L^2$ norm of the change in shape between iterations
\item Difference in reactivity between iterations
\item Difference in amplitude between iterations
\item IQS uniqueness consistency criteria
\end{enumerate}
\begin{itemize}
\item 1-4 had equivalent behavior
\item 5 required analytical integration of precursors for reasonable error convergence
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Conclusions (cont.)}

\begin{block}{IQS Time Step Error Convergence}
\begin{itemize}
\item Demonstrated proper convergence through order 4
\item Required higher order interpolation of PRKE parameters for higher order convergence
\item Effectiveness of error convergence demonstrated through time adaptation
\item Marginal improvement over implicit discretization for "difficult" 1D problem
\item Significant improvement for TWIGL and LRA
\end{itemize}
\end{block}

\begin{block}{IQS with Temperature Feedback}
\begin{itemize}
\item IQS was successfully implemented into the Rattlesnake/MOOSE framework
\item Added intermediate time scale loop for temperature
\item Incorporation of temperature into the quasi-static process significantly improved computational efficiency
\item Dynamical time scale analysis demonstrated proof-of-concept for variant time scales
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Recommendations for Future Research}

\begin{block}{Multiphysics}
Test IQS with other reactor physics, including:
\begin{itemize}
\item More complex temperature feedback
\item Thermal Hydraulics
\item Fuel Mechanics
\end{itemize}
\end{block}

\begin{block}{Transport}
Apply IQS to other transport models (other Rattlesnake action systems):
\begin{itemize}
\item DFEM-Diffusion
\item SAAF-Sn
\item SAAF-Pn
\item LS-Sn
\end{itemize}
\end{block}

\begin{block}{PJFNK}
\begin{itemize}
\item Investigate PJFNK iteration of IQS
\item Optimize IQS preconditioning
\end{itemize}
\end{block}

@@

%-------------------------------------------------------------------
\begin{frame}{Fixed-point iteration}
%\vspace{-10mm}

\begin{figure}
\includegraphics[height=1.3in]{figures/fixed_point.png}
\end{figure}
%\vspace{-10mm}
\begin{block}{IQS iteration approach}
\setlength{\leftmargini}{10mm}
\begin{itemize}
\item[\textit{Step 1:}] Compute the PRKE parameters at the end of the macro step using the last computed shape
\item[\textit{Step 2:}] Linearly interpolate the computed PRKE parameters over the macro step
\item[\textit{Step 3:}] Solve the PRKE on micro steps over the entire macro step
\item[\textit{Step 4:}] Solve the shape equation on the macro step using the computed values of $p$ and $dp/dt$.
\item[\textit{Step 5:}] Check if the shape solution has converged:
	\begin{itemize}
	\item \textit{No:} Repeat the same macro time step
	\item \textit{Yes:} Move on to the next macro time step
	\end{itemize}
\end{itemize}
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Fixed-point iteration programming logic}

\begin{figure}[!htpb]
\centering
\resizebox{!}{3in}{
\begin{tikzpicture}[every node/.style = {font=\LARGE}]

\node[greenblock](p1) at (0,0) {Solve shape equations \\ over macro time step};
\node[greenblock](p2) at (0,-2.5) {Compute PRKE parameters \\ over macro time step};
\node[greenblock](p3) at (0,-5) {Solve PRKE using micro \\ steps};
\node[greenblock](p4) at (0,-7.5) {Update $p$ and $\frac{dp}{dt}$};
\node[blueblock2] (p5) at (0,-11) {Solve other physics \\ components};
\node[reddiamond] (check) at (0,-15) {Check for \\ convergence};
\node [above =0.1mm of p1] {\large{IQS Solve:}};
\node [above =0.1mm of p5] {\large{Multi-physics:}};

\tikzback{p1}{p1}{p4}{p4}{bk1}
\tikzback{p5}{p5}{p5}{p5}{bk2}

\draw[->,ultra thick](p1.south) -- (p2.north);
\draw[->,ultra thick](p2.south) -- (p3.north);
\draw[->,ultra thick](p3.south) -- (p4.north);
\draw [->,ultra thick] (p4.south)+(0,-0.25) -- node [above] {} (bk2-n);
\draw[->,ultra thick](p5.south)+(0,-0.25) -- (check.north);
%\draw[->,ultra thick](check.west) -- (-5,-15) -- node[above,sloped] {\large{no}}(-5,0) -- node[above] {}(bk1-w);
\draw[->,ultra thick](check.west) -- node[above,sloped] {\LARGE{no}} (-5,-15) |-  node[above] {}(bk1-w);
\draw[->,ultra thick](check.south) -- node[right] {\LARGE{yes}} ++(0,-2);

%-------------------------------------------------------------------
\begin{frame}{Newton iteration}
\vspace{-3mm}
\begin{figure}
\includegraphics[height=1.1in]{figures/newton_iter.jpg}
\end{figure}
\vspace{-3mm}
\begin{block}{Jacobian-Free Newton-Krylov Method}
\bi
\item Newton system:
\be
J\delta\varphi = F(\varphi,p,t) - A(\varphi,p)\varphi \equiv -R(\varphi,p)
\ee
\item Where $J$ is the Jacobian and defined as:
\be
J_{ij}=\partial R_i/\partial\varphi_j
\ee
\item For IQS, the Jacobian is impossible to define so it must be approximated:
\be
J\delta\varphi\approx[R(\varphi+\epsilon\delta\varphi,p')-R(\varphi,p)]/\epsilon
\ee
\item The resulting system is solved using GMRES iteration
\ei
\end{block}

%-------------------------------------------------------------------
\begin{frame}{Time Discretization Schemes}
\vspace{-3mm}

\begin{block}{General Equation Form}
\be 
IV\frac{\partial \varphi}{\partial t} = A\varphi+b
\ee
\end{block}
\vspace{-3mm}

\begin{block}{Theta Method}
\be
\varphi_{n+1} = (IV-\Delta t A_{n+1})^{-1}\left[\Delta t(1-\theta)(A_n\varphi_n+b_n) + \Delta t\theta b_{n+1} + IV\varphi_n\right]
\ee
\begin{itemize}
\item Implicit Euler: $\theta=1$ (First Order)
\item Explicit Euler: $\theta=0$ (First Order)
\item Crank-Nicolson: $\theta=1/2$ (Second Order)
\end{itemize}
\end{block}
\vspace{-3mm}

>>>>

\setbeamertemplate{footline}{
\leavevmode%
%\hbox{\hspace*{-0.06cm}
\begin{beamercolorbox}[wd=.5\paperwidth,ht=3.25ex,dp=1ex,center]{author in head/foot}%
	\usebeamerfont{author in head/foot}\insertshortauthor%~~(\insertshortinstitute)
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.25\paperwidth,ht=3.25ex,dp=1ex,center]{section in head/foot}%
	\usebeamerfont{section in head/foot} M\&C 2017 Jeju, Korea % \insertshorttitle
\end{beamercolorbox}%
\begin{beamercolorbox}[wd=.25\paperwidth,ht=3.25ex,dp=1ex,left]{section in head/foot}%
	\usebeamerfont{section in head/foot}\insertshortdate{}\hspace*{2em}
	\insertframenumber{} / \inserttotalframenumber %\hspace*{2ex}
\end{beamercolorbox}}%
%\vskip0pt%
%}

@@

\title{Multiphysics Core Dynamics Simulation Using the Improved Quasi-Static Method}

\author[Zachary M. Prince]{\underline{Zachary M. Prince} and Jean C. Ragusa}
\institute{Department of Nuclear Engineering, Texas A\&M University, College Station, TX}

\logo{\includegraphics[height=0.8cm]{figures/IUP_Logo.jpeg}\includegraphics[height=0.8cm]{figures/TAM-logo.png}}

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

%-------------------------------------------------------------------
\begin{frame}{Conclusions}
\vspace{-3mm}

\begin{block}{Summary}
\begin{itemize}
\item Derivation of IQS and implementation of quasi-static time stepping
\item Semi-analytic evaluation of temperature
\item Implementation of adiabatic heat up in quasi-static process
\item Testing with LRA benchmark and TREAT Transient-15 model
\end{itemize}
\end{block}

\begin{block}{Conclusions}
\begin{itemize}
\item Temperature-amplitude iteration decreased temperature-shape iteration
\item Semi-analytic temperature evaluation improved IQS CPU performance
\item Intermediate time scale significantly improved IQS time step performance.
\item Optimal CPU performance with 1 update for IQS and 4 updates for IQS P-C
\item Time constant analysis showed adaptation could further improve performance
\end{itemize}
\end{block}

\begin{block}{Future Work}
\begin{itemize}
\item Implement intermediate time scale to more advanced multiphysics
\item Develop adaptation technique for number of multipysics updates
\end{itemize}
\end{block}

@@
